---
title: "New York School District Contingency Budgets: Revote Outcomes by Cap Status"
subtitle: "Built from NYSED statewide budget-vote and revote/defeated files"
author: "Prepared for Don Boyd"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true
    theme: cosmo
params:
  years: !expr 2012:2026     # tax-cap era; pre-2012 has no cap, so above/below-cap is undefined
  cache_dir: "data/raw/nysed_budget_votes"   # downloaded NYSED .xls/.xlsx live here
  download: false            # files already cached; set TRUE to (re)download from NYSED
  levy_tol: 0.0005           # tolerance (in proportion units) when testing proposed levy > levy limit
execute:
  warning: true
  message: true
---

```{r setup}
#| label: setup
#| include: true
suppressPackageStartupMessages({
  library(tidyverse) # dplyr, tidyr, stringr, purrr, readr, ggplot2
  library(readxl) # reads BOTH .xls and .xlsx
  library(janitor) # clean_names()
  library(gt) # presentation tables
})
have_httr2 <- requireNamespace("httr2", quietly = TRUE) # optional; download.file used if absent
if (have_httr2) {
  suppressPackageStartupMessages(library(httr2))
}

options(timeout = 120)
dir.create(params$cache_dir, showWarnings = FALSE, recursive = TRUE)
```

## 0. What this notebook does (and an honest caveat)

NYSED posts two Excel files per year on its *Statewide School District Budget Voting
Results* page (https://p12.nysed.gov/mgtserv/votingresults/):

1. **Budget Vote Results** — the first (May) vote, all districts.
2. **Budgets Defeated / Revotes** — the districts that failed the first vote, with their
   June revote result (if they held one).

A district operates on a **contingency budget** when it (a) fails the first vote and then
fails the revote, **or** (b) fails the first vote and chooses to skip the revote and adopt
contingency directly. The revote file is what lets us separate those cases.

**Caveat that drives the design.** NYSED changes the column headers and even the file
format (.xls vs .xlsx) from year to year, sometimes adding title rows above the table.
Because these files are binary and weren't openable in the environment where this notebook
was drafted, the column names below are matched by **keyword patterns**, not hard-coded
names, and Section 3 **prints the raw top rows and detected headers for every file** so you
can confirm/adjust before trusting the downstream tables. Treat Sections 4–7 as correct
*conditional on* the keyword map in Section 4 resolving to the right columns — verify that
once against the Section 3 output.

## 1. File registry (exact NYSED URLs)

```{r registry}
#| label: registry
# Exact URLs copied from the NYSED voting-results index. `skip` = number of title rows to
# drop before the header row; default 0, override after inspecting Section 3.
# Note: 2017 has no published revote file; pre-2012 omitted (no tax cap → no cap test).
registry <- tribble(
  ~year , ~kind    , ~url                                                                                        , ~skip ,
   2026 , "vote"   , "https://p12.nysed.gov/mgtserv/votingresults/2025-26-budget-vote-results.xlsx"              ,     0 ,
   2026 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/2025-26-budget-revote-results.xlsx"            ,     0 ,
   2025 , "vote"   , "https://p12.nysed.gov/mgtserv/votingresults/2024-25-budget-vote-results-final.xlsx"        ,     0 ,
   2025 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/2024-25-budget-revote-results-final.xlsx"      ,     0 ,
   2024 , "vote"   , "https://p12.nysed.gov/mgtserv/votingresults/2023-24_Budget_Vote_Results_FINAL.xlsx"        ,     0 ,
   2024 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/2023-24_Budget_Revote_Results_FINAL.xlsx"      ,     0 ,
   2023 , "vote"   , "https://p12.nysed.gov/mgtserv/votingresults/docs/2022-23_Budget_Vote_Results_FINAL.xlsx"   ,     0 ,
   2023 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/docs/2022-23-budget-revote-results.xlsx"       ,     0 ,
   2022 , "vote"   , "https://p12.nysed.gov/mgtserv/documents/2021-22-budget-vote-results_final.xlsx"            ,     0 ,
   2022 , "revote" , "https://p12.nysed.gov/mgtserv/documents/2021-22-budget-revote-results_final.xlsx"          ,     0 ,
   2021 , "vote"   , "https://www.p12.nysed.gov/mgtserv/documents/2020-21-budget-vote-results_post.xlsx"         ,     0 ,
   2021 , "revote" , "https://www.p12.nysed.gov/mgtserv/documents/2020-21_Budget_Re-Vote_Results_FINAL.xlsx"     ,     0 ,
   2020 , "vote"   , "https://www.p12.nysed.gov/mgtserv/votingresults/2019-20BudgetVoteResults_POST_FFinal.xlsx" ,     0 ,
   2020 , "revote" , "https://www.p12.nysed.gov/mgtserv/votingresults/SCHOOLBUDGETRE-VOTE6-18-19.xlsx"           ,     0 ,
   2019 , "vote"   , "https://p12.nysed.gov/mgtserv/documents/2018SchoolBudgetVoteResults_POST.xls"              ,     0 ,
   2019 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/2018SchoolBudgetRevoteResults_Final.xlsx"      ,     0 ,
   2018 , "vote"   , "https://p12.nysed.gov/mgtserv/documents/Copyof2017budgetvoteresults5-16-17.xls"            ,     0 ,
   # 2018 revote: none published
  2017 , "vote"   , "https://p12.nysed.gov/mgtserv/documents/2016budgetvoteresultsmay17_POST_rjune29.xls"       ,     0 ,
   2017 , "revote" , "https://p12.nysed.gov/mgtserv/documents/2016BudgetRevotes_Post_Final_R.xls"                ,     0 ,
   2016 , "vote"   , "https://p12.nysed.gov/mgtserv/votingresults/docs/2015budgetvoteresults.xls"                ,     0 ,
   2016 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/docs/2015-16budgetdefeat.xls"                  ,     0 ,
   2015 , "vote"   , "https://p12.nysed.gov/mgtserv/votingresults/docs/2014budgetvoteresults.xls"                ,     0 ,
   2015 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/docs/2014-15budgetdefeat.xls"                  ,     0 ,
   2014 , "vote"   , "https://p12.nysed.gov/mgtserv/votingresults/docs/2013budgetvoteresults.xls"                ,     0 ,
   2014 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/docs/2013-14budgetdefeat.xls"                  ,     0 ,
   2013 , "vote"   , "https://p12.nysed.gov/mgtserv/votingresults/docs/2012budgetvoteresults.xls"                ,     0 ,
   2013 , "revote" , "https://p12.nysed.gov/mgtserv/votingresults/docs/2012budgetdefeat-1.xls"                   ,     0
) |>
  # NB: the `year` literals here are the budget-END year (NYSED labels files "2024-25" →
  # 2025); the download chunk uses them for filenames, then converts `year` to the vote's
  # calendar year (May 2024 vote → 2024). `params$years` is matched on the budget-end label.
  filter(year %in% params$years)

registry |> count(kind)
```

## 2. Download & cache

```{r download}
#| label: download
fetch_one <- function(url, dest) {
  if (file.exists(dest) && !params$download) {
    return(dest)
  }
  ok <- TRUE
  if (have_httr2) {
    tryCatch(
      {
        req <- httr2::request(url) |>
          httr2::req_timeout(120) |>
          httr2::req_user_agent(
            "contingency-budget-analysis (educational; contact: you@example.org)"
          )
        resp <- httr2::req_perform(req)
        writeBin(httr2::resp_body_raw(resp), dest)
      },
      error = function(e) {
        tryCatch(
          download.file(url, dest, mode = "wb", quiet = TRUE),
          error = function(e2) {
            ok <<- FALSE
            warning("FAILED: ", url, " :: ", conditionMessage(e2))
          }
        )
      }
    )
  } else {
    tryCatch(
      download.file(url, dest, mode = "wb", quiet = TRUE),
      error = function(e2) {
        ok <<- FALSE
        warning("FAILED: ", url, " :: ", conditionMessage(e2))
      }
    )
  }
  if (!file.exists(dest) || file.size(dest) < 1000) {
    ok <- FALSE
    warning("Downloaded file looks empty/blocked: ", url)
  }
  if (ok) dest else NA_character_
}

registry <- registry |>
  mutate(
    ext = if_else(str_detect(url, "\\.xlsx$"), ".xlsx", ".xls"),
    # Build the cached path from the budget-END year (NYSED's "2025-26" -> 2026, which is
    # how the files are named on disk) ...
    dest = file.path(params$cache_dir, sprintf("%d_%s%s", year, kind, ext)),
    dest = map2_chr(url, dest, fetch_one),
    # ... THEN relabel `year` to the vote's CALENDAR year (the May-2025 vote -> 2025). Vote
    # dates embedded in the files confirm this (the "2026" file's vote date is 2025-05-20).
    # This is the convention used everywhere below and by the Section 7 cross-checks.
    year = year - 1L
  )

registry |> select(year, kind, ext, dest) |> mutate(ok = !is.na(dest))
```

> If many rows show `ok = FALSE`, you're likely on a network that blocks `p12.nysed.gov`
> (the drafting sandbox was). Run this where the site is reachable, or download the files
> by hand into `params$cache_dir` using the naming `"<year>_<vote|revote>.<ext>"`.

## 3. Inspect raw structure (DO THIS BEFORE trusting later sections)

```{r inspect}
#| label: inspect
#| results: asis
peek <- function(dest, year, kind) {
  if (is.na(dest) || !file.exists(dest)) {
    return(invisible(NULL))
  }
  cat(sprintf("\n\n### %d — %s\n\n", year, kind))
  raw_head <- tryCatch(
    read_excel(dest, col_names = FALSE, n_max = 6, .name_repair = "minimal"),
    error = function(e) NULL
  )
  if (!is.null(raw_head)) {
    cat("**First 6 rows (no header assumed):**\n\n")
    print(raw_head)
  }
  this_skip <- registry$skip[registry$year == year & registry$kind == kind][1]
  hdr <- tryCatch(
    names(read_excel(
      dest,
      skip = this_skip,
      n_max = 1,
      .name_repair = "unique_quiet"
    )),
    error = function(e) character(0)
  )
  cat("\n**Detected header names (after `skip`):**\n\n")
  cat(paste0("- `", janitor::make_clean_names(hdr), "`"), sep = "\n")
}
pwalk(list(registry$dest, registry$year, registry$kind), peek)
```

## 4. Layout detection + harmonization

Each file is reduced to a standard schema. NYSED sheets are *not* tidy tables — they carry
1–2 title rows, **multi-row merged headers**, sometimes a leading index column or BEDS code,
and (crucially) **no levy data at all**, only a "60% required" supermajority flag. So instead
of matching a single cleaned header row, `detect_layout()` infers each sheet's structure:

- **district name** — the column whose header text contains `"SCHOOL DISTRICT"`.
- **Yes / No counts** — the adjacent numeric pair whose `yes/(yes+no)` best matches the file's
  own `"% Yes"` column; failing that, the `Total = Yes + No` triple (preferring the *revote*
  block in revote files and never the *additional propositions* block); last resort, the two
  right-most numeric columns.
- **60% flag** — the column of `Y/N` values (`TRUE` = above cap, 60% needed).

This was validated against all 27 files: where a `"% Yes"` column exists the detected Yes/No
reconcile with it exactly, and computed pass/defeat counts match NYSED's own `P`/`D` markers.

```{r harmonize}
#| label: harmonize
# Coerce a possibly-formatted numeric cell ("1,248", "75.1%") to a plain number.
num <- function(x) suppressWarnings(as.numeric(gsub("[,$%]", "", as.character(x))))
is_numlike <- function(x) !is.na(num(x))
`%or%` <- function(a, b) if (length(a)) a else b # first if non-empty, else second

# Locate name / Yes / No / 60%-flag columns and the first data row in one raw
# (header-less, all-text) sheet. See Section 4 prose for the detection rules.
detect_layout <- function(raw, kind) {
  nc <- ncol(raw)
  nr <- nrow(raw)

  # Header block = the top rows (within the first 10) carrying any header token.
  toks <- c(
    "\\bYes\\b", "\\bNo\\b", "Total", "BUDGET", "60%", "Approval",
    "\\(Y/N\\)", "% *Yes", "PROPOSITION", "REVOTE", "FIRST"
  )
  hdr_rows <- which(vapply(seq_len(min(nr, 10)), function(i) {
    any(vapply(toks, function(t) any(grepl(t, raw[i, ], ignore.case = TRUE)), logical(1)))
  }, logical(1)))
  header_end <- if (length(hdr_rows)) max(hdr_rows) else 0L
  data_idx <- (header_end + 1):nr
  hdr <- raw[seq_len(max(header_end, 1)), , drop = FALSE]

  # District name: column whose header text contains "SCHOOL DISTRICT".
  name_col <- which(vapply(seq_len(nc), function(j) {
    any(grepl("SCHOOL DISTRICT", hdr[, j], ignore.case = TRUE))
  }, logical(1)))
  name_col <- if (length(name_col)) {
    name_col[1]
  } else {
    ft <- vapply(seq_len(nc), function(j) {
      mean(!is_numlike(raw[data_idx, j]) & nchar(trimws(raw[data_idx, j])) > 4, na.rm = TRUE)
    }, numeric(1))
    which.max(ft)
  }

  # Column type profiles over NON-BLANK data cells (blank region rows dilute otherwise).
  frac <- function(j, pred) {
    v <- raw[data_idx, j]
    nb <- !is.na(v) & trimws(v) != ""
    if (sum(nb) < 3) 0 else mean(pred(v[nb]))
  }
  cnt <- function(j, pred) {
    v <- raw[data_idx, j]
    nb <- !is.na(v) & trimws(v) != ""
    sum(nb & pred(v))
  }
  numcols <- which(
    vapply(seq_len(nc), frac, numeric(1), pred = is_numlike) > 0.7 &
      vapply(seq_len(nc), cnt, numeric(1), pred = is_numlike) >= 3
  )
  numcols <- numcols[numcols > name_col] # vote counts sit to the right of the name

  # "% Yes" share column -- anchored at the cell start so "60% YES NEEDED" is excluded.
  share_col <- which(vapply(seq_len(nc), function(j) {
    any(grepl("^\\s*%\\s*Yes", hdr[, j], ignore.case = TRUE))
  }, logical(1)))
  share_col <- if (length(share_col)) share_col[1] else NA_integer_

  yes_col <- no_col <- NA_integer_

  # (1) Adjacent numeric pair whose share best matches the "% Yes" column.
  if (!is.na(share_col)) {
    sv <- num(raw[data_idx, share_col])
    mv <- median(sv, na.rm = TRUE)
    if (!is.na(mv) && mv > 1.5) sv <- sv / 100 # stored as percent, not proportion
    best <- NULL
    best_err <- Inf
    for (a in numcols) {
      for (b in numcols[numcols > a & numcols <= a + 2]) {
        sh <- num(raw[data_idx, a]) / (num(raw[data_idx, a]) + num(raw[data_idx, b]))
        err <- median(abs(sh - sv), na.rm = TRUE)
        if (isTRUE(is.finite(err)) && err < best_err) {
          best_err <- err
          best <- c(a, b)
        }
      }
    }
    if (!is.null(best) && best_err < 0.02) {
      yes_col <- best[1]
      no_col <- best[2]
    }
  }

  # (2) Total = Yes + No triple, choosing the right block by its header label.
  if (is.na(yes_col)) {
    triples <- list()
    for (a in numcols) {
      b <- a + 1L
      cc <- a + 2L
      if (b %in% numcols && cc %in% numcols) {
        ok <- num(raw[data_idx, a]) + num(raw[data_idx, b]) == num(raw[data_idx, cc])
        if (mean(ok, na.rm = TRUE) > 0.8 && sum(!is.na(ok)) > 3) {
          lbl <- ""
          for (jj in a:1) {
            hit <- hdr[, jj][grepl(
              "BUDGET|REVOTE|PROPOSITION|FIRST|ADDITIONAL", hdr[, jj],
              ignore.case = TRUE
            )]
            if (length(hit)) {
              lbl <- toupper(hit[1])
              break
            }
          }
          triples[[length(triples) + 1]] <- list(a = a, b = b, lbl = lbl)
        }
      }
    }
    if (length(triples)) {
      if (kind == "revote") {
        rv <- Filter(function(t) grepl("REVOTE", t$lbl), triples)
        pick <- if (length(rv)) {
          rv[[1]]
        } else {
          (Filter(function(t) !grepl("PROPOSITION|ADDITIONAL", t$lbl), triples) %or% triples)[[1]]
        }
      } else {
        nb <- Filter(function(t) !grepl("PROPOSITION|ADDITIONAL|REVOTE", t$lbl), triples)
        pick <- (if (length(nb)) nb else triples)[[1]]
      }
      yes_col <- pick$a
      no_col <- pick$b
    }
  }

  # (3) Last resort: the two right-most numeric columns.
  if (is.na(yes_col) && length(numcols) >= 2) {
    l2 <- tail(numcols, 2)
    yes_col <- l2[1]
    no_col <- l2[2]
  }

  # 60% supermajority flag: a column of Y/N values.
  ynf <- vapply(seq_len(nc), frac, numeric(1), pred = function(v) toupper(trimws(v)) %in% c("Y", "N"))
  ync <- vapply(seq_len(nc), cnt, numeric(1), pred = function(v) toupper(trimws(v)) %in% c("Y", "N"))
  cand <- which(ynf > 0.6 & ync >= 3)
  cand <- cand[cand > name_col]
  flag_col <- if (length(cand)) cand[1] else NA_integer_

  list(
    name_col = name_col, yes_col = yes_col, no_col = no_col,
    flag_col = flag_col, data_idx = data_idx
  )
}

harmonize <- function(dest, year, kind) {
  if (is.na(dest) || !file.exists(dest)) {
    return(NULL)
  }
  raw <- as.data.frame(
    read_excel(dest, col_names = FALSE, col_types = "text", .name_repair = "minimal"),
    stringsAsFactors = FALSE
  )
  lay <- detect_layout(raw, kind)
  if (is.na(lay$yes_col) || is.na(lay$no_col)) {
    warning(sprintf("[%d-%s] could not locate Yes/No columns", year, kind))
    return(NULL)
  }
  di <- lay$data_idx
  district <- str_squish(raw[di, lay$name_col])
  yes <- num(raw[di, lay$yes_col])
  no <- num(raw[di, lay$no_col])
  flag <- if (!is.na(lay$flag_col)) {
    toupper(trimws(raw[di, lay$flag_col])) == "Y"
  } else {
    rep(NA, length(di))
  }
  # Drop title/section/total rows and anything without an actual vote.
  keep <- !is.na(district) & district != "" &
    !str_detect(district, regex("statewide|percentage|override distric|^boces", ignore_case = TRUE)) &
    !is.na(yes) & !is.na(no) & (yes + no) > 0

  tibble(
    year,
    kind,
    district_id = NA_character_, # BEDS codes absent / inconsistent across file generations
    district = district[keep],
    county = NA_character_, # county not carried in these files
    yes = yes[keep],
    no = no[keep],
    levy_pct = NA_real_, # NYSED vote files contain NO levy data ...
    levy_limit_pct = NA_real_, # ... so cap status keys off override_flag (below)
    override_flag = flag[keep] # TRUE = needed 60% to pass = proposed levy above the cap
  )
}

clean_all <- registry |>
  mutate(data = pmap(list(dest, year, kind), harmonize))

votes <- clean_all |>
  filter(kind == "vote") |>
  pull(data) |>
  compact() |>
  list_rbind()
revotes <- clean_all |>
  filter(kind == "revote") |>
  pull(data) |>
  compact() |>
  list_rbind()

glimpse(votes)
glimpse(revotes)
```

## 5. Classify cap status and outcomes

```{r classify}
#| label: classify
# Override (above cap) test: explicit flag if present, else proposed levy > levy limit.
needs_override <- function(levy_pct, limit_pct, flag, tol = params$levy_tol) {
  case_when(
    !is.na(flag) ~ flag,
    !is.na(levy_pct) & !is.na(limit_pct) ~ levy_pct > (limit_pct + tol),
    TRUE ~ NA
  )
}
# Pass test depends on threshold: override needs >=60% yes; otherwise simple majority.
passed_vote <- function(yes, no, is_override) {
  share <- yes / (yes + no)
  thr <- if_else(is_override %in% TRUE, 0.60, 0.50)
  case_when(
    is.na(share) ~ NA,
    is_override %in% TRUE ~ share >= thr,
    TRUE ~ share > thr
  )
}

v1 <- votes |>
  transmute(
    year,
    district_id,
    district,
    county,
    yes1 = yes,
    no1 = no,
    levy_pct1 = levy_pct,
    levy_limit_pct1 = levy_limit_pct,
    override1 = needs_override(levy_pct, levy_limit_pct, override_flag),
    yes_share1 = yes / (yes + no),
    passed1 = passed_vote(yes, no, override1)
  )

v2 <- revotes |>
  transmute(
    year,
    district_id,
    district,
    county,
    yes2 = yes,
    no2 = no,
    levy_pct2 = levy_pct,
    levy_limit_pct2 = levy_limit_pct,
    override2 = needs_override(levy_pct, levy_limit_pct, override_flag),
    yes_share2 = yes / (yes + no),
    passed2 = passed_vote(yes, no, override2),
    revote_held = !is.na(yes) & !is.na(no) & (yes + no) > 0
  )

# Join revote info onto the set of districts that failed the first vote.
# Prefer BEDS id; fall back to (year, district) name match if ids are missing.
join_keys <- if (
  mean(!is.na(v1$district_id)) > 0.9 && mean(!is.na(v2$district_id)) > 0.9
) {
  c("year", "district_id")
} else {
  c("year", "district")
}
message("Joining vote↔revote on: ", paste(join_keys, collapse = ", "))

failed_first <- v1 |> filter(passed1 %in% FALSE)

# Keep ONLY the join keys plus revote-specific measures, so we never drop a key we're
# joining on and never collide with v1's district/county columns.
v2_join <- v2 |>
  select(
    all_of(join_keys),
    yes2,
    no2,
    levy_pct2,
    levy_limit_pct2,
    override2,
    yes_share2,
    passed2,
    revote_held
  )

panel <- failed_first |>
  left_join(v2_join, by = join_keys) |>
  mutate(
    revote_held = coalesce(revote_held, FALSE),
    # Cap status governing the SECOND event (what Don asked to split on):
    #  - if a revote was held, use override2 (the revote budget's cap status)
    #  - if no revote (direct contingency), fall back to override1
    second_override = coalesce(override2, override1),
    cap_status = case_when(
      second_override %in% TRUE ~ "Above cap (override / 60%)",
      second_override %in% FALSE ~ "At/below cap (simple majority)",
      TRUE ~ "Unknown cap status"
    ),
    outcome = case_when(
      revote_held & passed2 %in% TRUE ~ "Passed on revote",
      revote_held & passed2 %in% FALSE ~ "Contingency (failed revote)",
      !revote_held ~ "Contingency (no revote / direct)",
      TRUE ~ "Unresolved (check data)"
    ),
    on_contingency = str_starts(outcome, "Contingency")
  )

glimpse(panel)
```

## 6. Summary tables

### 6a. Contingency budgets per year

```{r tbl-contingency-per-year}
#| label: tbl-contingency-per-year
per_year <- panel |>
  group_by(year) |>
  summarise(
    failed_first_vote = n(),
    held_revote = sum(revote_held),
    passed_on_revote = sum(outcome == "Passed on revote"),
    contingency_total = sum(on_contingency),
    .groups = "drop"
  ) |>
  arrange(year)

per_year |>
  gt() |>
  tab_header(
    title = "First-vote defeats, revotes, and contingency budgets by year",
    subtitle = "Source: NYSED statewide vote & revote files (computed)"
  ) |>
  cols_label(
    failed_first_vote = "Failed 1st vote",
    held_revote = "Held revote",
    passed_on_revote = "Passed on revote",
    contingency_total = "On contingency"
  ) |>
  grand_summary_rows(
    columns = c(
      failed_first_vote,
      held_revote,
      passed_on_revote,
      contingency_total
    ),
    fns = list(Total ~ sum(.))
  )
```

### 6b. Revote outcomes by cap status — passed vs. contingency

This is the core table you asked for: for second votes **above the cap** and **at/below the
cap**, how many passed vs. went to contingency. (Faceted by year; collapse with the second
chunk for an all-years view.)

```{r tbl-cap-outcome}
#| label: tbl-cap-outcome
cap_outcome <- panel |>
  mutate(
    outcome2 = if_else(
      on_contingency,
      "Went to contingency",
      "Passed on revote"
    )
  ) |>
  count(year, cap_status, outcome2) |>
  pivot_wider(names_from = outcome2, values_from = n, values_fill = 0) |>
  arrange(year, cap_status)

cap_outcome |>
  gt(groupname_col = "year") |>
  tab_header(title = "Revote outcomes by cap status, by year")

# All-years pooled
panel |>
  mutate(
    outcome2 = if_else(
      on_contingency,
      "Went to contingency",
      "Passed on revote"
    )
  ) |>
  count(cap_status, outcome2) |>
  pivot_wider(names_from = outcome2, values_from = n, values_fill = 0) |>
  gt() |>
  tab_header(
    title = sprintf(
      "Revote outcomes by cap status — pooled %d–%d",
      min(params$years),
      max(params$years)
    )
  )
```

### 6c. Per-district revote detail

Every district that failed its first vote, with its cap status (60%-required flag), both vote
%s, and the outcome — sortable/paged. (NYSED files carry no levy figures, so there are no
levy-% columns here.)

```{r tbl-district-detail}
#| label: tbl-district-detail
detail <- panel |>
  transmute(
    year,
    district,
    vote1_pct_for = round(100 * yes_share1, 1),
    vote1_pct_against = round(100 * (1 - yes_share1), 1),
    vote2_pct_for = round(100 * yes_share2, 1),
    vote2_pct_against = round(100 * (1 - yes_share2), 1),
    above_cap_1st = override1, # needed 60% on the first vote
    above_cap_2nd = second_override, # cap status governing the second event
    outcome
  ) |>
  arrange(year, desc(str_starts(outcome, "Contingency")), district)

# Show as an interactive paged frame; write a CSV for your own slicing.
dir.create("output", showWarnings = FALSE)
readr::write_csv(detail, file.path("output", "revote_district_detail.csv"))
detail
```

## 7. Sanity checks against independently verified values

Hard-coded facts from secondary reporting (NYSSBA/Empire Center/OSC press releases and
district clerks). Treat these as **ballpark cross-checks, not exact targets**: these sources
count "defeats" and "override budgets" on conventions that differ from the raw NYSED files
(e.g. NYSSBA's 2025 "23 defeats / 42 override budgets" vs. NYSED's own `P`/`D` and 60%-flag
markers, which this pipeline reproduces exactly at 19 and 34). A *large* gap, or a flipped
sign, is the signal worth chasing in Section 4 — small differences are expected.

```{r known-values}
#| label: known-values
known_year_level <- tribble(
  ~year , ~fact                                                                         , ~value , ~source                                                 ,
   2012 , "Districts adopting contingency after 1st vote (first tax-cap year)"          ,      0 , "NYSSBA Budget Defeats & Contingency Budgets 1992-2012" ,
   2020 , "Districts on contingency (Rensselaer, Riverhead, Fort Edward + East Ramapo)" ,      4 , "OSC/DiNapoli 2020 budget-vote report"                  ,
   2024 , "First-vote defeats"                                                          ,     18 , "NYSSBA press release 2024-05-22"                       ,
   2024 , "Override budgets sought / failed on 1st vote"                                ,     10 , "NYSSBA press release 2024-05-22 (10 of 32 failed)"     ,
   2025 , "First-vote defeats"                                                          ,     23 , "NYSSBA press release 2025-05-21 (651 approved of 674)" ,
   2025 , "Override budgets sought / failed on 1st vote"                                ,     17 , "NYSSBA press release 2025-05-21 (17 of 42 failed)"     ,
   2025 , "Revoting districts (override-failures / simple-majority-failures = 17 / 7)"  ,     24 , "Empire Center 2025-06-16"
)
known_year_level |> gt() |> tab_header(title = "Known year-level checkpoints")

# District-level checkpoints (fully specified worked examples) — verified from district clerks
# and local reporting. Levy figures are % change in the tax LEVY (not spending).
known_district <- tribble(
  ~year , ~district              , ~cap_pct , ~v1_levy_pct , ~v2_levy_pct , ~v1_for , ~v1_ag , ~v2_for , ~v2_ag , ~v1_override , ~v2_override , ~outcome                           ,
   2025 , "Mahopac CSD"          , NA       , 0.010        , 0.000        , 48.8    , 51.2   , 71.0    , 29.0   , FALSE        , FALSE        , "Passed on revote"                 ,
   2025 , "Elwood UFSD"          , 0.0129   , 0.0420       , 0.0264       , 54.2    , 45.8   , 71.6    , 28.4   , TRUE         , TRUE         , "Passed on revote"                 ,
   2025 , "Corning-Painted Post" , NA       , NA           , NA           , 40.3    , 59.7   , 58.1    , 41.9   , TRUE         , FALSE        , "Passed on revote"                 ,
   2024 , "Corning-Painted Post" , NA       , 0.049        , 0.0297       , 50.2    , 49.8   , 67.0    , 33.0   , TRUE         , FALSE        , "Passed on revote"                 ,
   2024 , "Ithaca CSD"           , NA       , 0.0842       , NA           , 29.5    , 70.5   , NA      , NA     , TRUE         , NA           , "Revote outcome not captured here" ,
   2020 , "Riverhead CSD"        , NA       , NA           , 0.0027       , NA      , NA     , 49.3    , 50.7   , FALSE        , FALSE        , "Contingency (failed revote)"
)
known_district |>
  gt() |>
  tab_header(
    title = "Known district-level checkpoints (verified worked examples)",
    subtitle = "Levy = % change in tax levy; blanks = not cleanly reported in public sources"
  ) |>
  fmt_percent(
    columns = c(cap_pct, v1_levy_pct, v2_levy_pct),
    decimals = 2,
    drop_trailing_zeros = FALSE
  )
```

## 8. Notes, gotchas, and extensions

- **Cap status comes from the 60% flag, not a levy comparison.** The NYSED vote/revote files
  contain **no levy figures** — no proposed-levy %, no tax-levy-limit %. The only cap signal
  they carry is the "Budget Required 60% for Approval (Y/N)" / "60% YES NEEDED TO PASS" flag,
  which NYSED sets precisely when a district's proposed levy exceeds its cap. So `override_flag`
  (above cap) = that flag, and `passed_vote()` applies the 60% threshold when it is `TRUE`,
  a simple majority otherwise. The `levy_pct*` columns are kept in the schema but are `NA`.
- **If you want the actual levy %s**, they live in a *different* source (the OSC property-tax-cap
  filings / NYSED tax-report data), keyed by district and year. Joining those in would let you
  reconstruct the literal *proposed levy > limit* test; `needs_override()` already prefers a
  real levy comparison whenever `levy_pct`/`levy_limit_pct` are populated.
- **Direct-to-contingency districts** (failed once, skipped the revote) appear in `panel`
  with `revote_held = FALSE`. East Ramapo (2020) is the canonical case. If a year's revote
  file *excludes* such districts, they'll be missing entirely — cross-check `failed_first`
  count against the NYSSBA defeat count for that year.
- **2020 was atypical**: all-absentee voting, shifted June revote date; turnout and dynamics
  aren't comparable to other years.
- **Year convention**: this notebook indexes by the *vote's* calendar year. NYSED filenames
  use the *budget* year ("2024-25"). Don't double-count when merging with other series.
- **Pre-2012**: omitted because there was no tax cap, so the above/below-cap split is
  undefined. The longer defeat/contingency *counts* (1992–2012) are available from NYSSBA's
  "Budget Defeats and Adopted Contingency Budgets" table if you want to extend Section 6a
  backward (no cap split).
- **Reproducibility**: `embed-resources: true` self-contains the HTML; the per-district CSV
  is written to `output/`. Source `.xls/.xlsx` are cached in `params$cache_dir`
  (`data/raw/nysed_budget_votes`); set `download: true` to refresh them from NYSED. Pin versions
  with `renv::snapshot()` if archiving.

```{r session}
#| label: session
sessionInfo()
```
