---
title: "New York School District Contingency Budgets: Revote Outcomes by Cap Status"
subtitle: "Built from NYSED statewide budget-vote and revote/defeated files"
author: "Prepared for Don Boyd"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true
    theme: cosmo
params:
  years: !expr 2012:2026     # tax-cap era; pre-2012 has no cap, so above/below-cap is undefined
  cache_dir: "nysed_cache"
  download: true             # set FALSE to re-run on already-cached files
  levy_tol: 0.0005           # tolerance (in proportion units) when testing proposed levy > levy limit
execute:
  warning: true
  message: true
---

```{r setup}
#| label: setup
#| include: true
suppressPackageStartupMessages({
  library(tidyverse)   # dplyr, tidyr, stringr, purrr, readr, ggplot2
  library(readxl)      # reads BOTH .xls and .xlsx
  library(janitor)     # clean_names()
  library(gt)          # presentation tables
})
have_httr2 <- requireNamespace("httr2", quietly = TRUE)  # optional; download.file used if absent
if (have_httr2) suppressPackageStartupMessages(library(httr2))

options(timeout = 120)
dir.create(params$cache_dir, showWarnings = FALSE, recursive = TRUE)
```

## 0. What this notebook does (and an honest caveat)

NYSED posts two Excel files per year on its *Statewide School District Budget Voting
Results* page (https://p12.nysed.gov/mgtserv/votingresults/):

1. **Budget Vote Results** — the first (May) vote, all districts.
2. **Budgets Defeated / Revotes** — the districts that failed the first vote, with their
   June revote result (if they held one).

A district operates on a **contingency budget** when it (a) fails the first vote and then
fails the revote, **or** (b) fails the first vote and chooses to skip the revote and adopt
contingency directly. The revote file is what lets us separate those cases.

**Caveat that drives the design.** NYSED changes the column headers and even the file
format (.xls vs .xlsx) from year to year, sometimes adding title rows above the table.
Because these files are binary and weren't openable in the environment where this notebook
was drafted, the column names below are matched by **keyword patterns**, not hard-coded
names, and Section 3 **prints the raw top rows and detected headers for every file** so you
can confirm/adjust before trusting the downstream tables. Treat Sections 4–7 as correct
*conditional on* the keyword map in Section 4 resolving to the right columns — verify that
once against the Section 3 output.

## 1. File registry (exact NYSED URLs)

```{r registry}
#| label: registry
# Exact URLs copied from the NYSED voting-results index. `skip` = number of title rows to
# drop before the header row; default 0, override after inspecting Section 3.
# Note: 2017 has no published revote file; pre-2012 omitted (no tax cap → no cap test).
registry <- tribble(
  ~year, ~kind,    ~url,                                                                                              ~skip,
  2026,  "vote",   "https://p12.nysed.gov/mgtserv/votingresults/2025-26-budget-vote-results.xlsx",                     0,
  2026,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/2025-26-budget-revote-results.xlsx",                   0,
  2025,  "vote",   "https://p12.nysed.gov/mgtserv/votingresults/2024-25-budget-vote-results-final.xlsx",               0,
  2025,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/2024-25-budget-revote-results-final.xlsx",             0,
  2024,  "vote",   "https://p12.nysed.gov/mgtserv/votingresults/2023-24_Budget_Vote_Results_FINAL.xlsx",               0,
  2024,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/2023-24_Budget_Revote_Results_FINAL.xlsx",             0,
  2023,  "vote",   "https://p12.nysed.gov/mgtserv/votingresults/docs/2022-23_Budget_Vote_Results_FINAL.xlsx",          0,
  2023,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/docs/2022-23-budget-revote-results.xlsx",              0,
  2022,  "vote",   "https://p12.nysed.gov/mgtserv/documents/2021-22-budget-vote-results_final.xlsx",                   0,
  2022,  "revote", "https://p12.nysed.gov/mgtserv/documents/2021-22-budget-revote-results_final.xlsx",                 0,
  2021,  "vote",   "https://www.p12.nysed.gov/mgtserv/documents/2020-21-budget-vote-results_post.xlsx",                0,
  2021,  "revote", "https://www.p12.nysed.gov/mgtserv/documents/2020-21_Budget_Re-Vote_Results_FINAL.xlsx",            0,
  2020,  "vote",   "https://www.p12.nysed.gov/mgtserv/votingresults/2019-20BudgetVoteResults_POST_FFinal.xlsx",        0,
  2020,  "revote", "https://www.p12.nysed.gov/mgtserv/votingresults/SCHOOLBUDGETRE-VOTE6-18-19.xlsx",                  0,
  2019,  "vote",   "https://p12.nysed.gov/mgtserv/documents/2018SchoolBudgetVoteResults_POST.xls",                     0,
  2019,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/2018SchoolBudgetRevoteResults_Final.xlsx",             0,
  2018,  "vote",   "https://p12.nysed.gov/mgtserv/documents/Copyof2017budgetvoteresults5-16-17.xls",                   0,
  # 2018 revote: none published
  2017,  "vote",   "https://p12.nysed.gov/mgtserv/documents/2016budgetvoteresultsmay17_POST_rjune29.xls",              0,
  2017,  "revote", "https://p12.nysed.gov/mgtserv/documents/2016BudgetRevotes_Post_Final_R.xls",                       0,
  2016,  "vote",   "https://p12.nysed.gov/mgtserv/votingresults/docs/2015budgetvoteresults.xls",                       0,
  2016,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/docs/2015-16budgetdefeat.xls",                         0,
  2015,  "vote",   "https://p12.nysed.gov/mgtserv/votingresults/docs/2014budgetvoteresults.xls",                       0,
  2015,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/docs/2014-15budgetdefeat.xls",                         0,
  2014,  "vote",   "https://p12.nysed.gov/mgtserv/votingresults/docs/2013budgetvoteresults.xls",                       0,
  2014,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/docs/2013-14budgetdefeat.xls",                         0,
  2013,  "vote",   "https://p12.nysed.gov/mgtserv/votingresults/docs/2012budgetvoteresults.xls",                       0,
  2013,  "revote", "https://p12.nysed.gov/mgtserv/votingresults/docs/2012budgetdefeat-1.xls",                          0
) |>
  # NB: NYSED labels files by *budget* year (e.g. "2024-25"); here `year` = the vote's calendar
  # year (May 2024 → 2024). Adjust the mapping above if you prefer the budget-year convention.
  filter(year %in% params$years)

registry |> count(kind)
```

## 2. Download & cache

```{r download}
#| label: download
fetch_one <- function(url, dest) {
  if (file.exists(dest) && !params$download) return(dest)
  ok <- TRUE
  if (have_httr2) {
    tryCatch(
      {
        req <- httr2::request(url) |> httr2::req_timeout(120) |>
          httr2::req_user_agent("contingency-budget-analysis (educational; contact: you@example.org)")
        resp <- httr2::req_perform(req)
        writeBin(httr2::resp_body_raw(resp), dest)
      },
      error = function(e) {
        tryCatch(
          download.file(url, dest, mode = "wb", quiet = TRUE),
          error = function(e2) { ok <<- FALSE; warning("FAILED: ", url, " :: ", conditionMessage(e2)) }
        )
      }
    )
  } else {
    tryCatch(
      download.file(url, dest, mode = "wb", quiet = TRUE),
      error = function(e2) { ok <<- FALSE; warning("FAILED: ", url, " :: ", conditionMessage(e2)) }
    )
  }
  if (!file.exists(dest) || file.size(dest) < 1000) {
    ok <- FALSE
    warning("Downloaded file looks empty/blocked: ", url)
  }
  if (ok) dest else NA_character_
}

registry <- registry |>
  mutate(
    ext  = if_else(str_detect(url, "\\.xlsx$"), ".xlsx", ".xls"),
    dest = file.path(params$cache_dir, sprintf("%d_%s%s", year, kind, ext)),
    dest = map2_chr(url, dest, fetch_one)
  )

registry |> select(year, kind, ext, dest) |> mutate(ok = !is.na(dest))
```

> If many rows show `ok = FALSE`, you're likely on a network that blocks `p12.nysed.gov`
> (the drafting sandbox was). Run this where the site is reachable, or download the files
> by hand into `params$cache_dir` using the naming `"<year>_<vote|revote>.<ext>"`.

## 3. Inspect raw structure (DO THIS BEFORE trusting later sections)

```{r inspect}
#| label: inspect
#| results: asis
peek <- function(dest, year, kind) {
  if (is.na(dest) || !file.exists(dest)) return(invisible(NULL))
  cat(sprintf("\n\n### %d — %s\n\n", year, kind))
  raw_head <- tryCatch(
    read_excel(dest, col_names = FALSE, n_max = 6, .name_repair = "minimal"),
    error = function(e) NULL
  )
  if (!is.null(raw_head)) { cat("**First 6 rows (no header assumed):**\n\n"); print(raw_head) }
  this_skip <- registry$skip[registry$year == year & registry$kind == kind][1]
  hdr <- tryCatch(
    names(read_excel(dest, skip = this_skip, n_max = 1, .name_repair = "unique_quiet")),
    error = function(e) character(0)
  )
  cat("\n**Detected header names (after `skip`):**\n\n")
  cat(paste0("- `", janitor::make_clean_names(hdr), "`"), sep = "\n")
}
pwalk(list(registry$dest, registry$year, registry$kind), peek)
```

## 4. Keyword column map + harmonization

Each file is reduced to a standard schema. Columns are located by **keyword regex** (tried
in order) against `janitor::clean_names()` output. After reviewing Section 3, tweak the
pattern vectors here if any `pick_col()` warning fires or resolves to the wrong field.

```{r harmonize}
#| label: harmonize
# Find first column whose cleaned name matches any pattern (regex, case-insensitive).
pick_col <- function(nms, patterns, label, file_id, required = FALSE) {
  for (p in patterns) {
    hit <- nms[str_detect(nms, regex(p, ignore_case = TRUE))]
    if (length(hit) >= 1) return(hit[[1]])
  }
  if (required) warning(sprintf("[%s] No column for '%s' (patterns: %s)",
                                 file_id, label, paste(patterns, collapse = " | ")))
  NA_character_
}

# Coerce a vote-count or numeric field that may arrive as text ("1,248") or proportion.
as_num <- function(x) suppressWarnings(readr::parse_number(as.character(x)))

# Normalize a "levy change" field to PROPORTION units (0.042), handling values stored as
# percent (4.2) or proportion (0.042). Heuristic: |median| > 1.5 → treat as percent.
to_prop <- function(x) {
  v <- as_num(x)
  if (all(is.na(v))) return(v)
  if (stats::median(abs(v), na.rm = TRUE) > 1.5) v / 100 else v
}

read_clean <- function(dest, skip) {
  read_excel(dest, skip = skip, guess_max = 5000, .name_repair = "unique_quiet") |>
    clean_names()
}

harmonize <- function(dest, year, kind, skip) {
  if (is.na(dest) || !file.exists(dest)) return(NULL)
  df  <- read_clean(dest, skip)
  nms <- names(df)
  fid <- sprintf("%d-%s", year, kind)

  col_id     <- pick_col(nms, c("beds", "sed_code", "district_id", "lea"),               "district_id", fid)
  col_name   <- pick_col(nms, c("district_name", "^district$", "school_district", "name"),"district",    fid, required = TRUE)
  col_county <- pick_col(nms, c("county"),                                                "county",       fid)
  col_yes    <- pick_col(nms, c("^yes$", "yes_vote", "votes_yes", "in_favor", "^for$"),   "yes",          fid, required = TRUE)
  col_no     <- pick_col(nms, c("^no$", "no_vote", "votes_no", "against", "opposed"),     "no",           fid, required = TRUE)
  # Proposed levy CHANGE (%) for this vote:
  col_levy   <- pick_col(nms, c("proposed.*levy.*(change|increase|percent|pct)",
                                "levy.*(change|increase).*percent", "tax_levy_increase",
                                "percent_change.*levy", "levy_percent"),                  "proposed_levy_pct", fid)
  # Tax levy LIMIT / cap (%): allowable growth — the legal override threshold driver.
  col_limit  <- pick_col(nms, c("tax_levy_limit", "levy_limit", "allowable.*levy",
                                "cap.*percent", "permissible.*levy"),                     "levy_limit_pct",    fid)
  # Some files instead carry an explicit "supermajority/override required" flag:
  col_over   <- pick_col(nms, c("super_?majority", "override", "60_percent", "threshold"),"override_flag",     fid)

  tibble(
    year, kind,
    district_id   = if (!is.na(col_id))     as.character(df[[col_id]])     else NA_character_,
    district      = as.character(df[[col_name]]),
    county        = if (!is.na(col_county)) as.character(df[[col_county]]) else NA_character_,
    yes           = as_num(df[[col_yes]]),
    no            = as_num(df[[col_no]]),
    levy_pct      = if (!is.na(col_levy))  to_prop(df[[col_levy]])  else NA_real_,
    levy_limit_pct= if (!is.na(col_limit)) to_prop(df[[col_limit]]) else NA_real_,
    override_flag = if (!is.na(col_over)) {
                      f <- tolower(as.character(df[[col_over]]))
                      str_detect(f, "y|true|1|over|super")
                    } else NA
  ) |>
    mutate(
      district = str_squish(district),
      district_id = str_remove_all(district_id, "\\D")  # keep digits of BEDS code
    ) |>
    filter(!is.na(district), district != "", !is.na(yes) | !is.na(no))
}

clean_all <- registry |>
  mutate(data = pmap(list(dest, year, kind, skip), harmonize))

votes   <- clean_all |> filter(kind == "vote")   |> pull(data) |> compact() |> list_rbind()
revotes <- clean_all |> filter(kind == "revote") |> pull(data) |> compact() |> list_rbind()

glimpse(votes)
glimpse(revotes)
```

## 5. Classify cap status and outcomes

```{r classify}
#| label: classify
# Override (above cap) test: explicit flag if present, else proposed levy > levy limit.
needs_override <- function(levy_pct, limit_pct, flag, tol = params$levy_tol) {
  case_when(
    !is.na(flag)                              ~ flag,
    !is.na(levy_pct) & !is.na(limit_pct)      ~ levy_pct > (limit_pct + tol),
    TRUE                                       ~ NA
  )
}
# Pass test depends on threshold: override needs >=60% yes; otherwise simple majority.
passed_vote <- function(yes, no, is_override) {
  share <- yes / (yes + no)
  thr   <- if_else(is_override %in% TRUE, 0.60, 0.50)
  case_when(is.na(share) ~ NA, is_override %in% TRUE ~ share >= thr, TRUE ~ share > thr)
}

v1 <- votes |>
  transmute(
    year, district_id, district, county,
    yes1 = yes, no1 = no, levy_pct1 = levy_pct,
    levy_limit_pct1 = levy_limit_pct,
    override1 = needs_override(levy_pct, levy_limit_pct, override_flag),
    yes_share1 = yes / (yes + no),
    passed1    = passed_vote(yes, no, override1)
  )

v2 <- revotes |>
  transmute(
    year, district_id, district, county,
    yes2 = yes, no2 = no, levy_pct2 = levy_pct,
    levy_limit_pct2 = levy_limit_pct,
    override2 = needs_override(levy_pct, levy_limit_pct, override_flag),
    yes_share2 = yes / (yes + no),
    passed2    = passed_vote(yes, no, override2),
    revote_held = !is.na(yes) & !is.na(no) & (yes + no) > 0
  )

# Join revote info onto the set of districts that failed the first vote.
# Prefer BEDS id; fall back to (year, district) name match if ids are missing.
join_keys <- if (mean(!is.na(v1$district_id)) > 0.9 && mean(!is.na(v2$district_id)) > 0.9) {
  c("year", "district_id")
} else {
  c("year", "district")
}
message("Joining vote↔revote on: ", paste(join_keys, collapse = ", "))

failed_first <- v1 |> filter(passed1 %in% FALSE)

# Keep ONLY the join keys plus revote-specific measures, so we never drop a key we're
# joining on and never collide with v1's district/county columns.
v2_join <- v2 |>
  select(all_of(join_keys),
         yes2, no2, levy_pct2, levy_limit_pct2, override2, yes_share2, passed2, revote_held)

panel <- failed_first |>
  left_join(v2_join, by = join_keys) |>
  mutate(
    revote_held = coalesce(revote_held, FALSE),
    # Cap status governing the SECOND event (what Don asked to split on):
    #  - if a revote was held, use override2 (the revote budget's cap status)
    #  - if no revote (direct contingency), fall back to override1
    second_override = coalesce(override2, override1),
    cap_status = case_when(
      second_override %in% TRUE  ~ "Above cap (override / 60%)",
      second_override %in% FALSE ~ "At/below cap (simple majority)",
      TRUE                        ~ "Unknown cap status"
    ),
    outcome = case_when(
      revote_held & passed2 %in% TRUE  ~ "Passed on revote",
      revote_held & passed2 %in% FALSE ~ "Contingency (failed revote)",
      !revote_held                      ~ "Contingency (no revote / direct)",
      TRUE                              ~ "Unresolved (check data)"
    ),
    on_contingency = str_starts(outcome, "Contingency")
  )

glimpse(panel)
```

## 6. Summary tables

### 6a. Contingency budgets per year

```{r tbl-contingency-per-year}
#| label: tbl-contingency-per-year
per_year <- panel |>
  group_by(year) |>
  summarise(
    failed_first_vote = n(),
    held_revote       = sum(revote_held),
    passed_on_revote  = sum(outcome == "Passed on revote"),
    contingency_total = sum(on_contingency),
    .groups = "drop"
  ) |>
  arrange(year)

per_year |>
  gt() |>
  tab_header(title = "First-vote defeats, revotes, and contingency budgets by year",
             subtitle = "Source: NYSED statewide vote & revote files (computed)") |>
  cols_label(failed_first_vote = "Failed 1st vote", held_revote = "Held revote",
             passed_on_revote = "Passed on revote", contingency_total = "On contingency") |>
  grand_summary_rows(columns = c(failed_first_vote, held_revote, passed_on_revote, contingency_total),
                     fns = list(Total ~ sum(.)))
```

### 6b. Revote outcomes by cap status — passed vs. contingency

This is the core table you asked for: for second votes **above the cap** and **at/below the
cap**, how many passed vs. went to contingency. (Faceted by year; collapse with the second
chunk for an all-years view.)

```{r tbl-cap-outcome}
#| label: tbl-cap-outcome
cap_outcome <- panel |>
  mutate(outcome2 = if_else(on_contingency, "Went to contingency", "Passed on revote")) |>
  count(year, cap_status, outcome2) |>
  pivot_wider(names_from = outcome2, values_from = n, values_fill = 0) |>
  arrange(year, cap_status)

cap_outcome |>
  gt(groupname_col = "year") |>
  tab_header(title = "Revote outcomes by cap status, by year")

# All-years pooled
panel |>
  mutate(outcome2 = if_else(on_contingency, "Went to contingency", "Passed on revote")) |>
  count(cap_status, outcome2) |>
  pivot_wider(names_from = outcome2, values_from = n, values_fill = 0) |>
  gt() |>
  tab_header(title = sprintf("Revote outcomes by cap status — pooled %d–%d",
                             min(params$years), max(params$years)))
```

### 6c. Per-district revote detail

Every district that failed its first vote, with the cap, both levy %s, both vote %s, and the
outcome — sortable/paged.

```{r tbl-district-detail}
#| label: tbl-district-detail
detail <- panel |>
  transmute(
    year, district, county,
    cap_levy_limit_pct = levy_limit_pct1,
    vote1_levy_pct = levy_pct1,
    vote2_levy_pct = levy_pct2,
    vote1_pct_for     = round(100 * yes_share1, 1),
    vote1_pct_against = round(100 * (1 - yes_share1), 1),
    vote2_pct_for     = round(100 * yes_share2, 1),
    vote2_pct_against = round(100 * (1 - yes_share2), 1),
    override_1st = override1, override_2nd = second_override,
    outcome
  ) |>
  arrange(year, desc(str_starts(outcome, "Contingency")), district)

# Show as an interactive paged frame; write a CSV for your own slicing.
readr::write_csv(detail, file.path(params$cache_dir, "revote_district_detail.csv"))
detail
```

## 7. Sanity checks against independently verified values

Hard-coded facts from primary reporting (NYSED/OSC/NYSSBA/Empire Center and district
clerks). After you run the pipeline, the computed numbers above should reproduce these. Any
mismatch usually means a column was mapped wrong in Section 4 (re-check Section 3).

```{r known-values}
#| label: known-values
known_year_level <- tribble(
  ~year, ~fact,                                                                         ~value, ~source,
  2012,  "Districts adopting contingency after 1st vote (first tax-cap year)",          0,      "NYSSBA Budget Defeats & Contingency Budgets 1992-2012",
  2020,  "Districts on contingency (Rensselaer, Riverhead, Fort Edward + East Ramapo)", 4,      "OSC/DiNapoli 2020 budget-vote report",
  2024,  "First-vote defeats",                                                          18,     "NYSSBA press release 2024-05-22",
  2024,  "Override budgets sought / failed on 1st vote",                                 10,     "NYSSBA press release 2024-05-22 (10 of 32 failed)",
  2025,  "First-vote defeats",                                                          23,     "NYSSBA press release 2025-05-21 (651 approved of 674)",
  2025,  "Override budgets sought / failed on 1st vote",                                 17,     "NYSSBA press release 2025-05-21 (17 of 42 failed)",
  2025,  "Revoting districts (override-failures / simple-majority-failures = 17 / 7)",   24,     "Empire Center 2025-06-16"
)
known_year_level |> gt() |> tab_header(title = "Known year-level checkpoints")

# District-level checkpoints (fully specified worked examples) — verified from district clerks
# and local reporting. Levy figures are % change in the tax LEVY (not spending).
known_district <- tribble(
  ~year, ~district,              ~cap_pct, ~v1_levy_pct, ~v2_levy_pct, ~v1_for, ~v1_ag, ~v2_for, ~v2_ag, ~v1_override, ~v2_override, ~outcome,
  2025,  "Mahopac CSD",          NA,       0.010,        0.000,        48.8,    51.2,   71.0,    29.0,   FALSE,        FALSE,        "Passed on revote",
  2025,  "Elwood UFSD",          0.0129,   0.0420,       0.0264,       54.2,    45.8,   71.6,    28.4,   TRUE,         TRUE,         "Passed on revote",
  2025,  "Corning-Painted Post", NA,       NA,           NA,           40.3,    59.7,   58.1,    41.9,   TRUE,         FALSE,        "Passed on revote",
  2024,  "Corning-Painted Post", NA,       0.049,        0.0297,       50.2,    49.8,   67.0,    33.0,   TRUE,         FALSE,        "Passed on revote",
  2024,  "Ithaca CSD",           NA,       0.0842,       NA,           29.5,    70.5,   NA,      NA,     TRUE,         NA,           "Revote outcome not captured here",
  2020,  "Riverhead CSD",        NA,       NA,           0.0027,       NA,      NA,     49.3,    50.7,   FALSE,        FALSE,        "Contingency (failed revote)"
)
known_district |>
  gt() |>
  tab_header(title = "Known district-level checkpoints (verified worked examples)",
             subtitle = "Levy = % change in tax levy; blanks = not cleanly reported in public sources") |>
  fmt_percent(columns = c(cap_pct, v1_levy_pct, v2_levy_pct), decimals = 2, drop_trailing_zeros = FALSE)
```

## 8. Notes, gotchas, and extensions

- **"Above cap" ≠ "above 2%."** The legal test is *proposed levy > the district's tax levy
  limit*, and that limit routinely exceeds 2% because of the formula's growth factor, PILOT
  changes, and the capital/exclusion adjustments. Central Square (2025) lost a near-50/50
  vote with a ~6.86% proposed levy — i.e., it was likely a *simple-majority* vote whose cap
  was high — which is exactly why this notebook keys off the levy-limit column rather than a
  flat 2% rule. Confirm the limit column resolves correctly in Section 4.
- **Levy vs. spending.** Local reporting frequently quotes the *spending* % change; only the
  *levy* % governs the cap and the contingency freeze. Section 4's `col_levy` patterns target
  levy fields — verify they aren't catching a spending column.
- **Direct-to-contingency districts** (failed once, skipped the revote) appear in `panel`
  with `revote_held = FALSE`. East Ramapo (2020) is the canonical case. If a year's revote
  file *excludes* such districts, they'll be missing entirely — cross-check `failed_first`
  count against the NYSSBA defeat count for that year.
- **2020 was atypical**: all-absentee voting, shifted June revote date; turnout and dynamics
  aren't comparable to other years.
- **Year convention**: this notebook indexes by the *vote's* calendar year. NYSED filenames
  use the *budget* year ("2024-25"). Don't double-count when merging with other series.
- **Pre-2012**: omitted because there was no tax cap, so the above/below-cap split is
  undefined. The longer defeat/contingency *counts* (1992–2012) are available from NYSSBA's
  "Budget Defeats and Adopted Contingency Budgets" table if you want to extend Section 6a
  backward (no cap split).
- **Reproducibility**: `embed-resources: true` self-contains the HTML; the per-district CSV
  is written to `params$cache_dir`. Pin versions with `renv::snapshot()` if archiving.

```{r session}
#| label: session
sessionInfo()
```
